The not so widening achievement gap: an interntational comparison of the evolution of the SES achievement gap from 2000-2015

### 2017-09-05 paper/paper.Rnw
Finished all tracking tables. Reorganizing next

final_avg_diff_table <-
  avg_diff_data %>%
  left_join(pred_avg_diff %>%
              filter(country %in% track_should_change & (num_tracks == 1 & age_selection == 1)) %>%
              select(country, pred), by = "country") %>%
  select(country, combn, avg_diff, pred) %>%
  separate(combn, c("num_tracks", "age_selection"), sep = "_") %>%
  transmute(
    Country = country,
    `Number of tracks` = ifelse(num_tracks == 0, "More than one track", "One track"),
    `Age of selection` = ifelse(age_selection == 1, "15 or more", "Less than 15"),
    `Actual inequality` = avg_diff %>% round(3),
    `Predicted inequality` = pred %>% round(3),
    `% reduction` = (((`Actual inequality` - `Predicted inequality`)/ `Actual inequality`) * 100) %>%
      ceiling %>% paste0("%")
    )

@@

<<>>=

gaps <- c("90th/10th SES gap", "80th/20th SES gap", "70th/30th SES gap")
all_gaps_models <-
  map(gaps, function(gap) {
    ready_data_age <-
      dif_data(complete_gaps, tracking_data, central_examination, which_gap = gap) %>%
      mutate(age_selection = ifelse(selage >= 15, 1, 0) %>% as.factor)

@@

# It under predicts the y value rather than over predict
# It has problems predicting the central tendency of the model
# This linear model explains the pooled 90/10 differences

ready_data <- dif_data(complete_gaps, tracking_data, central_examination)

# pp_check(mod_tracking, type = "dens_overlay", nsamples = 100)
# pp_check(mod_tracking, type = "scatter_avg", nsamples = 100)
# pp_check(mod_tracking, type = "boxplot", nsamples = 20)
# pp_check(mod_tracking, type = "error_hist", nsamples = 6)
# pp_check(mod_tracking, type = "ecdf_overlay")
# 
# pp_check(mod_tracking, type = "error_scatter_avg_vs_x", x = "zvoc")
# pp_check(mod_tracking, type = "error_scatter_avg_vs_x", x = "ztrack")

<<>>=

@@

ready_data_two <-
  ready_data %>%
  group_by(country) %>%
  summarize_at(vars(vars_unique), unique) %>%
  mutate(cum_diff = ready_data %>% group_by(country) %>% summarize(m = sum(difference)) %>% pull(m))

mod_tracking <-
  ready_data_two %>%
  lm(cum_diff ~ ztrack + zvoc, data = .) %>%
  arm::display(detail = T)

mod_tracking <-
  ready_data_two %>%
  filter(!is.na(age_selection)) %>%
  brm(
    cum_diff ~ I(age_selection >= 15),
    family = gaussian(),
    data = .,
    warmup = 1000, iter = 2000, chains = 5
  )

pp_check(mod_tracking, type = "dens_overlay", nsamples = 100)
pp_check(mod_tracking, type = "ecdf_overlay", nsamples = 10)
pp_check(mod_tracking, type = "boxplot", nsamples = 30)
pp_check(mod_tracking, type = "stat", nsamples = 200) # very nice to compare central tendency
pp_check(mod_tracking, type = "error_scatter_avg", nsamples = 1000)

pp_check(mod_tracking, type = "loo_pit")
# over prediction
# central tendency is not captured very well
# more under-prediction than over prediction

@@

m_data <-
  ready_data_two %>%
  filter(!is.na(age_selection)) %>%
  mutate(pred_median = posterior_predict(mod_tracking) %>% colMeans(),
         ratio_pred = pred_median/cum_diff) %>%
  select(country, ratio_pred, cum_diff, pred_median)

>>>>

final_avg_diff_table <-
  avg_diff_data %>%
  left_join(pred_avg_diff %>%
              filter(country %in% track_should_change & (num_tracks == 1 & age_selection == 1)) %>%
              select(country, pred), by = "country") %>%
  select(country, combn, avg_diff, pred) %>%
  separate(combn, c("num_tracks", "age_selection"), sep = "_") %>%
  transmute(
    Country = country,
    `Number of tracks` = ifelse(num_tracks == 0, "More than one track", "One track"),
    `Age of selection` = ifelse(age_selection == 1, "15 or more", "Less than 15"),
    `Inequality` = avg_diff %>% round(3),
    `Predicted` = pred %>% round(3),
    `% reduction` = (((`Inequality` - `Predicted`)/ `Inequality`) * 100) %>%
      ceiling %>% paste0("%")
    )

@@

<<interaction_model>>=
ready_data <- dif_data(complete_gaps, tracking_data, central_examination)

# pp_check(mod_tracking, type = "dens_overlay", nsamples = 100)
# pp_check(mod_tracking, type = "scatter_avg", nsamples = 100)
# pp_check(mod_tracking, type = "boxplot", nsamples = 20)
# pp_check(mod_tracking, type = "error_hist", nsamples = 6)
# pp_check(mod_tracking, type = "ecdf_overlay")
# 
# pp_check(mod_tracking, type = "error_scatter_avg_vs_x", x = "zvoc")
# pp_check(mod_tracking, type = "error_scatter_avg_vs_x", x = "ztrack")
@

<<all_gaps_plot>>=
gaps <- c("90th/10th SES gap", "80th/20th SES gap", "70th/30th SES gap")
all_gaps_models <-
  map(gaps, function(gap) {
    ready_data_age <-
      dif_data(complete_gaps, tracking_data, central_examination, which_gap = gap) %>%
      mutate(age_selection = ifelse(selage >= 15, 1, 0) %>% as.factor)

@@

# It under predicts the y value rather than over predict
# It has problems predicting the central tendency of the model
# This linear model explains the pooled 90/10 differences
@

<<cumulative_modeling_table>>=

@@

all_gaps_models_cum <-
  map(gaps, function(gap) {

    gap_data <- dif_data(complete_gaps, tracking_data, central_examination, which_gap = gap)

    ready_data_age <-
      gap_data %>%
      mutate(age_selection = ifelse(selage >= 15, 1, 0) %>% as.factor) %>%
      group_by(country) %>%
      summarize_at(vars(vars_unique), unique) %>%
      mutate(cum_diff =
               gap_data %>%
               group_by(country) %>%
               summarize(m = sum(difference)) %>%
               pull(m))

    mod_tracking <-
      ready_data_age %>%
      filter(!is.na(age_selection)) %>%
      brm(
        cum_diff ~ ztrack + zvoc,
        family = gaussian(),
        data = .,
        warmup = 1000, iter = 2000, chains = 5
      )

tidy_dfs <-
  map(all_gaps_models_cum, ~ {

  broom::tidy(.x) %>%
    .[grepl("^b_", .$term), ] %>%
    mutate_if(is_double, round, 2) %>%
    transmute(term,
              estimate = paste0(estimate, " (", lower, "/", upper, ")"))
  })

tidy_dfs_two <-
  reduce(tidy_dfs, full_join, by = "term") %>%
  setNames(c(" ", gsub(" SES gap", "", gaps))) %>%
  mutate(" " = c(
    "Intercept",
    "Track Index",
    "Vocational Index"
  )) %>%
  .[c(2, 3, 1), ]

r_square <-
  paste0("\\hline ", paste0("R square: & ",
         paste0(map(all_gaps_models_cum, ~ bayes_R2(.x)[1] %>% round(2) * 100) %>% paste0("\\%"),
                       collapse = " & ")), " \\\\")
n_obs <-
  paste0(paste0(" Sample size: & ",
         paste0(map(all_gaps_models_cum, nobs), collapse = " & ")), " \\\\ \\hline")

addtorow_two <- list(pos = list(pos = 3), command = command_to_row)

<<>>=
# pp_check(mod_tracking, type = "dens_overlay", nsamples = 100)
# pp_check(mod_tracking, type = "ecdf_overlay", nsamples = 10)
# pp_check(mod_tracking, type = "boxplot", nsamples = 30)
# pp_check(mod_tracking, type = "stat", nsamples = 200) # very nice to compare central tendency
# pp_check(mod_tracking, type = "error_scatter_avg", nsamples = 1000)
# 
# pp_check(mod_tracking, type = "loo_pit")
# over prediction
# central tendency is not captured very well
# more under-prediction than over prediction

@@

m_data <-
  ready_data_two %>%
  filter(!is.na(ztrack), !is.na(zvoc)) %>%
  mutate(pred_median = posterior_predict(all_gaps_models_cum[[1]]) %>% colMeans(),
         ratio_pred = pred_median/cum_diff) %>%
  select(country, ratio_pred, cum_diff, pred_median)
